---
title: "An Analysis of Early Career Data Scientists using Kaggle Data"
author: "John Dinh"
date: '2022-03-15'
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, message = F, echo = F, warning = F}
library(here)
library(naniar)
library(tidyverse)
library(wesanderson)
library(scales)
library(vtable)
library(ggplot2)
library(ggpubr)
library(skimr)
library(kableExtra)
library(knitr)
library(themis)
library(caret)
library(car)
library(randomForest)
library(pROC)
library(magrittr)
library(jtools)
library(finalfit)
library(ggthemes)
library(imbalance)
#Custom theme
theme_set(theme_tufte())
my_theme <- theme(plot.title = element_text(hjust = 0.5, face = 2, size = 18),
        plot.subtitle = element_text(hjust = 0.5, size = 13),
        axis.title = element_text(face = 1, size = 15),
        axis.text = element_text(size = 13))

# Custom palette
my_palette <- c('#FAD510', '#CB2314', '#273046', '#354823', '#1E1E1E',
                "#A42820", "#5F5647", "#9B110E", "#3F5151", "#4E2A1E")

```


```{r}
data <- read.csv(here('healthcare-dataset-stroke-data.csv'))
```
# Data Cleaning

First let's take a look at the overall data quality. There are no missing observations in this data.

```{r}
glimpse(data)
cols <- c('gender', 'ever_married', 'work_type', 'Residence_type','smoking_status', 'stroke', 'hypertension', 'heart_disease')
data %<>% mutate_at(cols, factor) %>% mutate(bmi = as.numeric(bmi))
```

There is a fairly large amount of unknown values for smoker status. 
```{r}
data %>% count(smoking_status) %>% rename(`Smoking Status` = smoking_status,
                                          Count = n) %>% 
  ggplot(aes(x = `Smoking Status`, y = Count))+
  geom_col(aes(Count, reorder(`Smoking Status`, Count)))


```

```{r}
#removing gender observation with other since it was only 1 observation
#replaced inconsistent values in BMI and smoking status with unknown 
data %<>% replace_with_na(data = data, replace = list(bmi = c('N/A'), smoking_status = c('Unknown'))) %>%  filter(gender != 'Other')
data1 %<>% replace_with_na(data = data, replace = list(bmi = c('N/A'), smoking_status = c('Unknown'))) %>%  filter(gender != 'Other')
```

## Imputation

The missing plot map shows there is no obvious patterns of missingness. We will impute values based on the assumption that values are missing completely at random (MCAR). 
```{r}
data %>% missing_plot()
```

Imputing Values for BMI. The plot below shows a slight right skew distribution for BMI. We can impute values for BMI using the median. The introduced randomness using imputation needs to be addressed. 
```{r, warning = F, echo = F}

bmi_preimute <- data1 %>% ggplot() + 
  geom_histogram(color = 'black', fill = my_palette[6], aes(x = bmi)) +
  labs(xaxis = 'BMI', title = 'BMI Distribution: Pre-Imputation') +
  xlab('BMI') +
  ylab('Count')

bmi_preimute
```

```{r}
#imputing median
data %<>% mutate(bmi = ifelse(is.na(bmi), median(bmi, na.rm = T),
                             bmi))
bmi_plt <- data %>% ggplot() + geom_histogram(color = 'black', fill = my_palette[6], aes(x = bmi)) +
  labs(xaxis = 'BMI', title = 'BMI Distribution: Median Imputation') +
  xlab('BMI') +
  ylab('Count')
bmi_plt
#smoking: imputing based on previous value since the missing plot show no pattern of missingness
data$smoking_status <- as.character(data$smoking_status)
data <- data %>% fill(smoking_status, .direction = 'downup') 

stroke_clean <- data %>% mutate(smoking_status = as.factor(smoking_status))
skim(data)
```


# Exploratory Data Analysis

There are more people that are older that get strokes. 
```{r}
age_density <- ggplot() + 
  geom_density(data=stroke_clean, aes(x=age , group=as.factor(stroke), fill=as.factor(stroke)), size=1,alpha=0.5, adjust=2)  + 
  ylab("Density")+ labs(fill=' ',x="Age")+
  scale_fill_manual(values = my_palette[2:3], labels = c('No stroke','Stroke')) +
  labs(title = 'Age vs Stroke Density Plot')

age_density
```


There is not much skewness in the stroke population compared to the non-stroke population. In the non-stroke population, the mean BMI is less than the median BMI. 

```{r}
bmi_density <- stroke_clean %>%  ggplot() + 
  geom_density( aes(x=bmi , group=as.factor(stroke), fill=as.factor(stroke)), size=1,alpha=0.5, adjust=2)  + 
  ylab("Density")+ labs(fill=' ',x="BMI")+
  scale_fill_manual(values = c(my_palette[1], my_palette[5]), labels = c('No stroke','Stroke')) +
  labs(title = 'BMI vs Stroke Density Plot')

bmi_density


```


```{r}
glucose_density <- stroke_clean %>%  ggplot() + 
  geom_density(aes(x=avg_glucose_level , group=as.factor(stroke), fill=as.factor(stroke)), size=1,alpha=0.5, adjust=2)  + 
  ylab("Density")+ labs(fill=' ', x= "Average Glucose Levels")+
  scale_fill_manual(values = c(my_palette[4], my_palette[5]), labels = c('No stroke','Stroke')) +
  labs(title = 'Average Glucose Levels vs Stroke Density Plot')
glucose_density

```
Having a stroke is associated with around 25 BMI and older population. 
```{r}

age_bmi <- stroke_clean %>% arrange(stroke) %>% 
  ggplot(aes(x = age, y = bmi, col = as.factor(stroke))) +
  geom_jitter(alpha = 2, aes(col = stroke))+
  scale_color_manual(labels = c('No stroke', 'Had a stroke'), values = c(my_palette[1], my_palette[5]))+
  labs(title = 'Scatter Plot: Age, BMI, Stroke', col = '') +
  xlab('Age')+
  ylab('BMI')

age_bmi

```

Stroke is also associated with older people here. The middle range of glucose levels have much fewer respondents of stroke however. 
```{r}
age_glucose_plt <- stroke_clean %>% arrange(stroke) %>% 
  ggplot(aes(x = age, y = avg_glucose_level, col = as.factor(stroke))) +
  geom_jitter(alpha = 2, aes(col = stroke))+
  scale_color_manual(labels = c('No stroke', 'Had a stroke'), values = c(my_palette[1], my_palette[9]))+
  labs(title = 'Scatter Plot: Age, Average Glucose Level, Stroke', col = '') +
  ylab('Average Glucose Level')+
  xlab('Age')

age_glucose_plt
```


```{r}
#stroke is heavily imbalanced
#pct <- stroke_clean %>% count(stroke) %>% mutate(pct = prop.table(n))
stroke_clean %>%  ggplot(aes(x = stroke)) +
  scale_x_discrete(labels = c('0' = 'Never had a stroke', '1'= 'Had a stroke'))+
  geom_bar(fill = my_palette[2]) +
  labs(x = '',
       y = 'Counts',
       title = 'Stroke Distribution')

```

## Train-Test Split

```{r}
set.seed(225231)
dt <- sort(sample(nrow(stroke_clean), nrow(stroke_clean) * .7))
train_stroke <- stroke_clean[dt,] %>% select(-id)
test_stroke <- stroke_clean[-dt,] %>% select(-id)
```


#Will use smote, oversampling technique since our dataset is not too large


```{r}
train_oversampled <- ovun.sample(stroke ~. , data = train_stroke, 
                                 method = 'over', 
                                 seed = 252323)
train_oversampled <- train_oversampled$data


```

Distribution of Stroke after Upsampling
```{r}
stroke_postsampling <- train_oversampled %>% ggplot(aes(x = stroke)) +
   scale_x_discrete(labels = c('0' = 'Never had a stroke', '1'= 'Had a stroke'))+
  geom_bar(fill = my_palette[3]) +
  labs(x = '',
       y = 'Counts',
       title = 'Stroke Distribution: Post-Upsampling')
stroke_postsampling

```
https://www.researchgate.net/publication/269177030_The_Impact_of_Oversampling_with_SMOTE_on_the_Performance_of_3_Classifiers_in_Prediction_of_Type_2_Diabetes

# Model Fitting

The ANOVA output shows a non significant chi-square value with p-value = .9. This means that the reduced model without the predictor,`bmi`, fits as well as the full model.  By the principle of parsimony, the more simple model will be used to assess the goodness of fit. 
```{r}
library(fastDummies)
train_oversampled_d <- train_oversampled
train_oversampled_d <- dummy_cols(train_oversampled_d, select_columns = c('gender', 'hypertension', 'heart_disease', 'ever_married','work_type', 'Residence_type',
                                                                          'smoking_status'),
                                  remove_selected_columns = T)
test_stroke_d <- test_stroke
test_stroke_d <- dummy_cols(test_stroke_d, select_columns = c('gender', 'hypertension', 'heart_disease', 'ever_married','work_type', 'Residence_type',
                                                                          'smoking_status'),
                                  remove_selected_columns = T)
 
full_fit <- glm(stroke ~ ., data = train_oversampled_d, family = 'binomial'(link = logit)) %>% tidy()
reduced_fit <- glm(stroke  ~ . , data = train_oversampled_d %>% select(-bmi), family = 'binomial'(link = logit)) %>%  tidy()

kable(anova(fit_reduced, fit_full, test = 'Chisq'))
```


```{r}
train_oversampled_d %>% 
  ggplot(aes())

```


```{r Pearson X^2 statistics} 
fit <- glm(stroke  ~ . , data = train_oversampled_d %>% select(-bmi), family = 'binomial'(link = logit))
sum(residuals(fit, type = "deviance")^2)/fit$df.residual
```


```{r}
probs <- fit %>% predict(test_stroke_d, type = 'response')
predicted_classes <- ifelse(probs > .5, '1', '0')
caret::confusionMatrix(data = as.factor(predicted_classes), reference = test_stroke_d$stroke)
```


```{r}
step <- step(glm(stroke~., data = train_oversampled, family ='binomial'(link = logit)),
             criterion = 'BIC')
summary(step)
```

```{r}
step_model <- glm(formula = stroke ~ gender + age + hypertension + heart_disease + 
    ever_married + work_type + Residence_type + avg_glucose_level + 
    smoking_status, family = binomial(link = logit), data = train_oversampled)

probs_step <- step_model %>% predict(test_stroke, type = 'response')
predicted_classes_step <- ifelse(probs_step > .5, '1', '0')
caret::confusionMatrix(data = as.factor(predicted_classes_step), reference = test_stroke$stroke)
```


```{r}
test_roc <- roc(test_stroke_d$stroke ~ probs, plot = T, print.auc = T)
```




